---
Time Series Analysis of COVID-19 Data Using R
---

#Load Required Libraries

```{r}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(janitor)
library(zoo)
```

#Import the Dataset

```{r}
covid19_data<-read.csv("X:/01StaffRoom/Mumthas/covid_19_data.csv", stringsAsFactors = FALSE)

covid19_data
```

#Check the structure:
```{r}
str(covid19_data)
```

#Display first 10 rows

```{r}
head(covid19_data, 10)
```

#Standardize column names

```{r}
covid19_data<- clean_names(covid19_data)
```

#DATA CLEANING

```{r}
covid19_data<- covid19_data %>% select(-last_update)
```

#Convert date column to date format

```{r}
covid19_data$observation_date<-as.Date(covid19_data$observation_date, format="%m/%d/%Y")
```

```{r}
covid19_data <- covid19_data %>%
  mutate(across(where(is.character), ~ na_if(.x, "")))
```


#Check Missing Values

```{r}
colSums(is.na(covid19_data))
```
#Replace missing province_state with unknown
```{r}
covid19_data<-covid19_data %>% mutate(province_state=replace_na(province_state,"unknown"))
```

#Duplicate Values

```{r}
sum(duplicated(covid19_data))
```

```{r}
covid19_data<- covid19_data %>% mutate(country_region= if_else(country_region=="Mainland China", "China", country_region))
```


#Month from observation date

```{r}
covid19_data<- covid19_data %>% mutate(month= month(observation_date, label = TRUE))
```


#How has the number of confirmed COVID-19 cases, deaths, and recoveries changed over time in a specific country?

```{r}
covid19_India <- covid19_data %>% filter(country_region=="India")
covid19_India
```

```{r}
cases_india<-covid19_India %>% group_by(observation_date) %>% 
  summarise( total_cases_confirmed = sum(confirmed),
             total_deaths_confirmed = sum(deaths),
             total_cases_recovered = sum(recovered),
             .groups = "drop")

cases_india
```


```{r}
ggplot(cases_india, aes(x=observation_date, y=total_cases_recovered))+
  geom_line(color="chocolate", linewidth = 1)+
  labs(title="Recovery rate in India",
       x="Date", y="Recovery Rate(%)")
```





#What is the trend of confirmed cases, deaths, and recoveries in India over time?

#Plot Confirmed Cases Trend

```{r}
ggplot(cases_india, aes(x=observation_date, y=total_cases_confirmed ))+
  geom_line(color="blue", linewidth=1)+
  scale_y_log10()+
  labs(title="COCID-19 confirmed cases trends in INDIA",
       x="Date", y="Number of confirmed cases")+
  theme_classic()
```

#Plot death Cases Trend

```{r}
ggplot(cases_india, aes(x=observation_date, y=total_deaths_confirmed ))+
  geom_line(color="blue", linewidth=1)+
 labs(title="COCID-19 deaths confirmed trends in INDIA",
       x="Date", y="Number of confirmed deaths")+
  theme_classic()
```

##Plot recovered  Cases Trend

```{r}
ggplot(cases_india, aes(x=observation_date, y=total_cases_recovered ))+
  geom_line(color="blue", linewidth=1)+
 labs(title="COCID-19 recovered  cases trends in INDIA",
       x="Date", y="Number of recovered cases")+
  theme_classic()
```

#Fatality & Recovery Rate Analysis

```{r}
covid19_India<- covid19_India %>% mutate(
  fatality_rate= deaths/confirmed*100,
  recovery_rate= recovered/confirmed*100
  )
```

#daily new COVID-19 cases change over time in India

```{r}
cases_india<- cases_india %>% 
  arrange(observation_date) %>%
  mutate(daily_new_cases = total_cases_confirmed -lag(total_cases_confirmed ,default=0) )
```

#Plot daily new cases in India

```{r}
ggplot(cases_india, aes(x=observation_date, y=daily_new_cases))+
  geom_line(color= "red", linewidth = 1)+
  labs(title="Daily New COVID-19 Cases in India",
       x="Date", y="Daily new cases")
  
  
```

#What is the smoothed trend of COVID-19 spread in India?

```{r}
cases_india <- cases_india %>%
  mutate(seven_day_average = rollmean(daily_new_cases,7, fill=NA))
```

# Plot 7 Day moving average

```{r}
ggplot(cases_india, aes(x=observation_date)) +
  geom_line(aes(y=daily_new_cases),alpha=0.05)+
  geom_line(aes(y=seven_day_average),color="blue", linewidth=1 )+
  labs(title="Seven Day Moving Averages of Daily new Covid19 Caes in India",
       x="date", y="Cases")
```


#How did COVID-19 cases evolve on a monthly basis in India?

```{r}
monthly_cases_india<- covid19_India %>%
  group_by(month)%>%
  summarize(total_confirmed= sum(confirmed),
            total_deaths= sum(deaths),
            total_recovered= sum(recovered) )
```

```{r}
ggplot(monthly_cases_india, aes(x=month, y=total_confirmed))+
   geom_col(fill = "steelblue") +
  labs(title="Monthly COVID-19 Confirmed Cases in India", 
       x="Month", y="cases")
```













